Ten Simple Rules for Teaching Programming

### 2017-07-04 paper.tex
A few small tidy-ups, adjusting from Markdown to LaTeX

Guzdial refers to the belief that some people are born programmers and others aren't
as ``computing's most enduring and damaging myth'' \citep{guzdial-myths}.
Our first and most important rule is that this is wrong:
competence at programming is not innate,
but is rather a learned skill that can be acquired and improved with practice.

@@

Beliefs such as this are known to have powerful effects on education outcomes [FIXME][fixme-outcomes].
If instructors believe that "some kids get it and some kids don't",
they will (consciously or unconsciously) invest less in those whom they put in the second category.
When combined with cultural stereotypes about who is and isn't a "natural programmer",
the downward spiral of under-achievement that results from differential attention
may be partly responsible for the gender imbalance in computing.

@@

\begin{enumerate}
\item It enables instructors to be more responsive to "what if?"
    questions. Where a slide deck is like a railway track, live coding
    allows instructors to go off road and follow their learners'
    interests.

\item  It facilitates unintended knowledge transfer: students learn more
    than the instructor consciously intends to teach by watching *how*
    instructors do things.

@@

\item  Typing in boilerplate code that is needed by the lesson, but not
    directly relevant to it (such as library import statements), can
    distract learners from the intended thrust of a lesson.  As
    \citet{willingham-book} says, "Memory is the residue of thought";
    if the instructor spends their time typing boilerplate, that may
    be all that learners take away.
\end{enumerate}

@@

1. The instructor gives learners a brief introduction to the topic.
1. The instructor then gives learners a multiple choice question
   that probes for misconceptions rather than simple factual recall.
1. Learners then vote on the answer to the question.
   * If they all have the right answer, move on.
   * If they all have the same wrong answer,
     the instructor addresses that specific misconception.
   * If they have a mix of right and wrong answers,
     they are given several minutes to discuss those answers with one another
     in small groups (typically 2-4 students)
     and then reconvene and vote again.

@@

For example,
a growing number of educators are including Parson's Problems in their pedagogic repertoire~\citep{parsons,morrison-parsons}.
Rather than writing programs from scratch,
learners are given the lines of code they need to solve a problem,
but in jumbled order.
Re-ordering them to solve the problem correctly allows them to concentrate on mastering control flow
without having to devote mental energy to recalling syntax
or the specifics of library functions.

>>>>

\citet{guzdial-myths} refers to the belief that some people are born programmers and others aren't
as ``computing's most enduring and damaging myth.''
Our first and most important rule is that this is wrong:
competence at programming is not innate,
but is rather a learned skill that can be acquired and improved with practice.

@@

Beliefs such as this are known to have powerful effects on education outcomes [FIXME][fixme-outcomes].
If instructors believe that ``some kids get it and some kids don't'',
they will (consciously or unconsciously) invest less in those whom they put in the second category.
When combined with cultural stereotypes about who is and isn't a "natural programmer",
the downward spiral of under-achievement that results from differential attention
may be partly responsible for the gender imbalance in computing.

@@

\begin{enumerate}
\item It enables instructors to be more responsive to ``what if?''
    questions. Where a slide deck is like a railway track, live coding
    allows instructors to go off road and follow their learners'
    interests.

\item  It facilitates unintended knowledge transfer: students learn more
    than the instructor consciously intends to teach by watching \textit{how}
    instructors do things.

@@

\item  Typing in boilerplate code that is needed by the lesson, but not
    directly relevant to it (such as library import statements), can
    distract learners from the intended thrust of a lesson.  As
    \citet{willingham-book} says, ``Memory is the residue of thought'';
    if the instructor spends their time typing boilerplate, that may
    be all that learners take away.
\end{enumerate}

@@

\begin{enumerate}
\item The instructor gives learners a brief introduction to the topic.
\item The instructor then gives learners a multiple choice question
   that probes for misconceptions rather than simple factual recall.
\item Learners then vote on the answer to the question.
  \begin{itemize}
   \item If they all have the right answer, move on.
   \item If they all have the same wrong answer,
     the instructor addresses that specific misconception.
   \item If they have a mix of right and wrong answers,
     they are given several minutes to discuss those answers with one another
     in small groups (typically 2-4 students)
     and then reconvene and vote again.
\end{itemize}
\end{enumerate}

@@

For example,
a growing number of educators are including Parsons Problems in their pedagogic repertoire~\citep{parsons,morrison-parsons}.
Rather than writing programs from scratch,
learners are given the lines of code they need to solve a problem,
but in jumbled order.
Re-ordering them to solve the problem correctly allows them to concentrate on mastering control flow
without having to devote mental energy to recalling syntax
or the specifics of library functions.
